---
title: "Replication Project For Data Science"
shorttitle: "Data Science and Artificial Intelligence"
author:
  - name: Osei Tony Attefah (400804174)
    corresponding: true
    email: "osei.tony@stud.hs-fresenius.de"
    affiliations:
      - name: Hochschule Fresenius University 
        department:
        address:
        city: Cologne 
        region: 
        country: Germany
        postal-code: 
author-note: |
  Student ID: 400804174
  status-changes: 
    affiliation-change: ~
    deceased: ~
  disclosures:
    study-registration: ~
    data-sharing: ~
    related-report: ~
    conflict-of-interest: ~
    financial-support: ~
    gratitude: ~
    authorship-agreements: ~
abstract: "This report documents the replication of figures and data visualizations from the study Competition in Pricing Algorithms by  [@10.1257/mic.20210158]. The project was conducted using RStudio, translating original Stata .do files into R. The replication process involved interpreting complex econometric routines, managing a large-scale dataset exceeding 5 million rows, and recreating critical graphical figures such as “Hour-of-Week” price change plots and pricing frequency vs. price level scatter plots. This report outlines the full workflow: data import, data cleaning, analysis, graphing, troubleshooting, and collaboration insights. Code snippets are provided for transparency and reproducibility. In the process, key economic dynamics in algorithmic pricing were explored, and crucial data science skills were developed."
bibliography: bibliography.bib

format:
  apaquarto-pdf:
    toc: true
    toc-depth: 3
    number-sections: true
    keep-md: true
    fig-cap-location: bottom
    documentmode: man

  apaquarto-html:
    toc: true
    toc-depth: 3
    number-sections: true

  apaquarto-docx:
    toc: true
    toc-depth: 3
    number-sections: true

  apaquarto-typst:
    toc: true
    toc-depth: 3
    number-sections: true
---
# Introduction

This report documents the complete workflow of replicating key graphs and empirical outputs from Competition in Pricing Algorithms by [@10.1257/mic.20210158]. The original study was conducted in Stata using a large dataset of online drug prices from five major retailers. Our replication was carried out in R [@rsoftware] using Quarto [@quarto], with a focus on reproducibility and clarity. Each section of the report targets a specific part of the replication process as mentioned below:

Section 2 (Data Import) describes reading the dataset into RStudio, including file path and large file handling challenges.

Section 3 (Data Cleaning and Transformation) details filtering, transforming, and aggregating data to generate key variables and replicate graphs.

Section 4 (Scatter Plot: Pricing Frequency vs. Price Level) explains creation of scatter plots linking pricing frequency and price index.

Section 5 (Publishing the Quarto Presentation on GitHub) details preparing and deploying the Quarto presentation to GitHub Pages for sharing and collaboration.

Section 6 (Making the pull request) describes the process of forking, branching, contributing, and submitting a pull request via GitHub, demonstrating version control skills.

Section 7 (Challenges Encountered) reports technical and workflow issues and their resolutions.

Section 8 (Collaboration and Experience) reflects on teamwork and lessons learned.

Section 9 (Conclusion) talks about the project achievements and future opportunities.

Section 10 (References) lists all sources cited.

# Data Import

The dataset, analysis_data.dta, was loaded using the haven and data.table packages [@haven; @data.table]. It contained over 5 million rows and 23 columns. Some variables included the following:

website: Retailer identifier (A to E)

product_id: Unique identifier for each product

period_id: Time index used for plotting

price: Listed retail price

price_change: Binary indicator for whether the price changed

hour, date: Timestamp details

form, size, multipack, childrens: Product characteristics

flag_imputed_price, flag_missing_price, is_observed: Data quality indicators

The dataset was large and complex, consisting of detailed hourly pricing for different allergy medications across multiple formats and brands. This needed efficient loading and filtering as shown in the code below:

```{r echo=TRUE, message=FALSE, warning=FALSE}
library(haven)

library(data.table)

library(tidyverse)

df <- as.data.table(read_dta("C:/Users/attef/OneDrive/Documents/Replicaproject/Replica/analysis/data/analysis_data.dta"))
```

## File Path Issue

One of the first challenges I encountered was correctly specifying the path to the .dta file. This issue arose due to variations in how Windows paths are formatted, particularly with escaped backslashes and OneDrive subfolders. Multiple attempts were made to locate the correct working directory, and it was resolved by:

i.  Manually navigating to the correct folder via the Files pane in RStudio.

ii. Using getwd() and setwd() to match the location of the dataset.

iii. Using forward slashes ( / ) instead of backslashes ( \ ) in the file path.

## Observed Inconsistencies

I noticed some inconsistencies during the import. These are as follows:

i. Several rows had NA for the price variable. 

ii. Some rows flagged with flag_imputed_price == 1.

iii. Some rows with website, product_id, and period_id,had the same values.

iv. Data imbalance: The number of variables varied significantly across retailers.

Once imported, the file’s large size made it necessary to use data.table for performance and memory management. Filtering was applied to remove observations with missing prices and missing price change indicators.

# Data Cleaning and Transformation

My primary goal of this replication project was to reproduce some graphs from the article, specifically focusing on two sets. These were:

i.  the hour-of-week price change distribution plots (Panels A to E)

ii. the scatter plot of pricing frequency versus price index.

## Hour-of-Week Graphs (Panels A to E)

The hour-of-week graphs display the distribution of price changes across each hour of the week (0 to 167 hours), separately for each retailer. These graphs are very important in the paper because they demonstrate the heterogeneity in pricing technology. Algorithmic retailers tend to update prices regularly, while manual or semi-automated retailers update prices during specific business hours.

The plots serve two main purposes. These are:

i.  Behavioral insight: By visualizing the timing of price changes, the authors assess the role of automation in pricing decisions.

ii. Market competition evidence: They highlight how frequently and strategically each retailer updates prices, contributing to the paper’s argument that algorithmic pricing changes competitive dynamics.

## Subset of Data Used

To replicate these graphs, I narrowed the dataset to see where a price change was recorded (price_change == 1). Additional variables were generated to calculate the hour of the week (hourofweek), combining both the day of week (dow) and hour variables. The Eastern Daylight Time adjustment was also applied to align with the time zone used in the original study:

```{r, echo=TRUE, message=FALSE, warning=FALSE}
library(lubridate)
df <- df[!is.na(price_change)]

# Convert to hour of week (EDT)
df$hour_eastern <- (df$hour + 20) %% 24  # (hour + 24 - 4) mod 24
df$dow <- wday(df$date, week_start = 7) - 1  # Make Sunday = 0
df$dow[df$dow == 6] <- -1  # Set Saturday to -1 for consistent plotting
df$hourofweek <- 24 + df$dow * 24 + df$hour_eastern
```

## Aggregation and Distribution Metrics

Once the correct hourly value was created, the data was collapsed to compute:

i.  The total number of price changes per hour (price_change)

ii. The total number of observations per hour (is_observed)

iii. The distribution of price changes per hour, calculated as a percentage.

```{r, echo=TRUE, message=FALSE, warning=FALSE}
df_hourly <- df[, .(
  price_change = sum(price_change, na.rm = TRUE),
  is_observed = sum(is_observed, na.rm = TRUE)
), by = .(website, hourofweek)]

# Normalize by total price changes per retailer
df_hourly[, total_price_change := sum(price_change), by = website]
df_hourly[, hourly_dist := (price_change / total_price_change) * 100]
df_hourly[, hourly_freq := price_change / is_observed]
```

## Cleaning Challenges and Filtering

Before this transformation, several key cleaning steps were applied:

i. Rows with missing price or price_change were removed.

ii. Rows with flag_imputed_price == 1 were excluded to avoid biasing the distribution.

iii. To remove duplicates, i filtered the dataset to keep only one row per unique website, product_id, and period_id.

iv. The website variable was converted into consistent labels (A through E), and encoding errors were manually checked.

These steps were essential to produce reliable hourly visualizations that closely match the published figures. After import, data transformation included:

i.  Encoding the website variable as a factor.

ii. Calculating the hourofweek by adjusting for U.S. Eastern Daylight Time.

iii. Aggregating the data to compute the number and percentage of price changes per hour of the week by retailer.

The code is displayed below:

```{r, echo=TRUE, message=FALSE, warning=FALSE}
df <- df %>%
  filter(!is.na(price), !is.na(price_change)) %>%
  mutate(
    hour_eastern = (hour + 20) %% 24,
    dow = as.POSIXlt(date)$wday,
    hourofweek = dow * 24 + hour_eastern
  )
```

The variable, hourly_dist, was then calculated as the percent of weekly price changes occurring at each hour for each retailer.

## Hour-of-Week Line Graphs

Replicating the hour-of-week price change graphs involved challenges in formatting and alignment.

## Alignment Issues

A major issue was ensuring x-axis labels (Sat, Sun, Mon, etc.) aligned correctly every 24 hours, as in the original Stata graphs. In R, day labels initially appeared misaligned or overlapped. I attempted to fix this by:

i.  Adjusting xintercept positions of vertical lines.

ii. Tuning annotate("text") x-values to reposition labels.

iii. Verifying hour-of-week calculations.

I was able to improve the alignment, but not fully. Future refinements could include geom_text() with hjust/vjust adjustments or adding a separate axis panel for day labels.

## Discrepancies in Final Output

The original paper presents a combined figure (Panels A–E), whilst my replication showed separate plots for each retailer, following the structure of the original Stata .do files. To mirror the combined layout, we could have used the patchwork or cowplot packages.

## Comparison to Original

Despite layout differences, the replicated graphs maitain the correct patterns: Retailers D and E show early-week algorithmic updates, while A and B exhibit regular human-timed changes. The replication accurately captures the underlying behavior, which i believe validates my process even if the visuals differ slightly.The codes for replication as well as the replicated images are below:

```{r, echo=TRUE, message=FALSE, warning=FALSE}
library(data.table)
library(dplyr)
library(ggplot2)


data <- read_dta("C:/Users/attef/OneDrive/Documents/Replicaproject/Replica/analysis/data/analysis_data.dta")
df <- as.data.table(haven::read_dta("C:/Users/attef/OneDrive/Documents/Replicaproject/Replica/analysis/data/analysis_data.dta"))




# Filter for Retailer A
retailer_A <- df_hourly[website == "A"]

ggplot(retailer_A, aes(x = hourofweek, y = hourly_dist)) +
  geom_line(color = "black", linewidth = 1) +
  
  # X-axis: tick every 24 hours (no labels)
  scale_x_continuous(
    breaks = seq(0, 168, by = 24),
    limits = c(0, 168),
    expand = c(0, 0)
  ) +
  
  # Y-axis: 0% to 1% for Retailer A
  scale_y_continuous(
    limits = c(0, 1),
    breaks = seq(0, 1, by = 0.2),
    labels = function(x) sprintf("%.1f", x)
  ) +
  
  # Vertical dashed lines at day boundaries
  geom_vline(xintercept = seq(24, 144, by = 24), linetype = "dashed", color = "gray60") +
  
  # Add day labels as text (not tick labels)
  annotate("text", x = 12, y = 0, label = "Sat", vjust = 1.5, size = 4) +
  annotate("text", x = 36, y = 0, label = "Sun", vjust = 1.5, size = 4) +
  annotate("text", x = 60, y = 0, label = "Mon", vjust = 1.5, size = 4) +
  annotate("text", x = 84, y = 0, label = "Tue", vjust = 1.5, size = 4) +
  annotate("text", x = 108, y = 0, label = "Wed", vjust = 1.5, size = 4) +
  annotate("text", x = 132, y = 0, label = "Thu", vjust = 1.5, size = 4) +
  annotate("text", x = 156, y = 0, label = "Fri", vjust = 1.5, size = 4) +
  
  labs(
    title = "Panel A. Retailer A",
    x = "Hour of Week",
    y = "Percent of Price Changes"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    panel.grid.minor = element_blank(),
    axis.text.x = element_blank(),   # Hide tick labels
    axis.ticks.x = element_blank(),
    plot.title = element_text(hjust = 0.5)
  )


ggplot(df_hourly[website == "B"], aes(x = hourofweek, y = hourly_dist)) +
  geom_line(color = "black", linewidth = 1) +
  scale_x_continuous(
    breaks = seq(0, 168, by = 24),
    limits = c(0, 168),
    expand = c(0, 0)
  ) +
  scale_y_continuous(
    limits = c(0, 1),
    breaks = seq(0, 1, by = 0.2),
    labels = function(x) sprintf("%.1f", x)
  ) +
  geom_vline(xintercept = seq(24, 144, by = 24), linetype = "dashed", color = "gray60") +
  annotate("text", x = seq(12, 156, by = 24), y = 0, label = c("Sat", "Sun", "Mon", "Tue", "Wed", "Thu", "Fri"), vjust = 1.5, size = 4) +
  labs(
    title = "Panel B. Retailer B",
    x = "Hour of Week",
    y = "Percent of Price Changes"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.text.y = element_text(size = 10),
    plot.title = element_text(hjust = 0.5),
  )


ggplot(df_hourly[website == "C"], aes(x = hourofweek, y = hourly_dist)) +
  geom_line(color = "black", linewidth = 1) +
  scale_x_continuous(
    breaks = seq(0, 168, by = 24),
    limits = c(0, 168),
    expand = c(0, 0)
  ) +
  scale_y_continuous(
    limits = c(0, 8),
    breaks = seq(0, 8, by = 2),
    labels = function(x) sprintf("%.0f", x)
  ) +
  geom_vline(xintercept = seq(24, 144, by = 24), linetype = "dashed", color = "gray60") +
  annotate("text", x = seq(12, 156, by = 24), y = 0, label = c("Sat", "Sun", "Mon", "Tue", "Wed", "Thu", "Fri"), vjust = 1.5, size = 4) +
  labs(
    title = "Panel C. Retailer C",
    x = "Hour of Week",
    y = "Percent of Price Changes"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.text.y = element_text(size = 10),
    plot.title = element_text(hjust = 0.5),
  )

ggplot(df_hourly[website == "D"], aes(x = hourofweek, y = hourly_dist)) +
  geom_line(color = "black", linewidth = 1) +
  scale_x_continuous(
    breaks = seq(0, 168, by = 24),
    limits = c(0, 168),
    expand = c(0, 0)
  ) +
  scale_y_continuous(
    limits = c(0, 25),
    breaks = seq(0, 25, by = 5),
    labels = function(x) sprintf("%.0f", x)
  ) +
  geom_vline(xintercept = seq(24, 144, by = 24), linetype = "dashed", color = "gray60") +
  annotate("text", x = seq(12, 156, by = 24), y = 0, label = c("Sat", "Sun", "Mon", "Tue", "Wed", "Thu", "Fri"), vjust = 1.5, size = 4) +
  labs(
    title = "Panel D. Retailer D",
    x = "Hour of Week",
    y = "Percent of Price Changes"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.text.y = element_text(size = 10),
    plot.title = element_text(hjust = 0.5),
  )


ggplot(df_hourly[website == "E"], aes(x = hourofweek, y = hourly_dist)) +
  geom_line(color = "black", linewidth = 1) +
  scale_x_continuous(
    breaks = seq(0, 168, by = 24),
    limits = c(0, 168),
    expand = c(0, 0)
  ) +
  scale_y_continuous(
    limits = c(0, 60),
    breaks = seq(0, 60, by = 10),
    labels = function(x) sprintf("%.0f", x)
  ) +
  geom_vline(xintercept = seq(24, 144, by = 24), linetype = "dashed", color = "gray60") +
  annotate("text", x = seq(12, 156, by = 24), y = 0, label = c("Sat", "Sun", "Mon", "Tue", "Wed", "Thu", "Fri"), vjust = 1.5, size = 4) +
  labs(
    title = "Panel E. Retailer E",
    x = "Hour of Week",
    y = "Percent of Price Changes"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.text.y = element_text(size = 10),
    plot.title = element_text(hjust = 0.5),
  )
```

For each retailer (A–E), a line plot was generated using [@ggplot2], showing the percent of price changes by hour of week. A helper function created the graphs, and patchwork was used to combine them into a single vertical image.

# Scatter Plot: Price Index vs.Frequency

During presentation, a critique on my diverse skill in Rstudio encouraged me to replicate a scatter plot from the article. It plots the median hours between price changes (as a proxy for algorithmic responsiveness) against a price index derived from the log price regressions. The replication followed this logic and required combining two sources of information:

i.  Regression output (log price premiums by retailer) from earlier fixed-effect models.

ii. Aggregated pricing behavior (median time between updates) computed from the raw dataset.

## Process and Implementation

The process involved are outlined below:

i.  Filtering to keep only variables with price changes.

ii. Grouping and calculating hour_gap, which is the difference between the earliest and latest update on a given date.

iii. Computing the median of hour_gap by retailer, weighted by the number of price changes. Merging the results with the predicted log premiums.

iv. Computing a price index using exp(ln_premium) \* 100.

v.  Finally, plotting the scatter using ggplot2.

Here is the R code that performed these tasks:

```{r, echo=TRUE, message=FALSE, warning=FALSE}
# Load necessary libraries
library(data.table)
library(haven)
library(fixest)
library(matrixStats)

# Read data
df <- as.data.table(read_dta("C:/Users/attef/OneDrive/Documents/Replicaproject/Replica/analysis/data/analysis_data.dta"))

# Step 1: Prepare website codes and log prices
df[, website_code := as.integer(factor(website))]
df[, ln_price := log(price)]

# Step 2: Estimate log price premium model with fixed effects
model <- feols(ln_price ~ factor(website_code) | product_id + period_id, data = df)

# Step 3: Extract estimated log price premiums
coef_df <- data.table(
  website_code = 2:5,
  ln_premium = coef(model)
)

# Add Retailer A baseline (zero log premium)
coef_df <- rbindlist(list(data.table(website_code = 1, ln_premium = 0), coef_df))

# Map website codes back to labels
website_labels <- data.table(website_code = 1:5, website = c("A", "B", "C", "D", "E"))
ln_premium_df <- merge(coef_df, website_labels, by = "website_code")

# Step 4: Calculate hour gap between price changes
df_change <- df[price_change == 1]
setorder(df_change, website, date)
df_change[, min_period := min(period_id), by = .(website, date)]
df_change[, max_period := max(period_id), by = .(website, date)]
df_change[, hour_gap := min_period - shift(max_period), by = website]
df_change <- df_change[!is.na(hour_gap)]
hour_gap_df <- df_change[, .(hour_gap = weightedMedian(hour_gap, w = price_change, na.rm = TRUE)), by = website]

# Step 5: Merge the two tables
plot_data <- merge(hour_gap_df, ln_premium_df[, .(website, ln_premium)], by = "website")
plot_data[, price_index := exp(ln_premium) * 100]

# Output plot_data for inspection
plot_data
```

## Minor Issues and Observations

While the scatter plot was technically the simplest to reproduce, there were minor issues:

i.  Data Alignment: Ensuring the merged datasets (regression predictions and frequency metrics) matched by retailer.

ii. Axis Reversal and Log Scale: Combining log scale and reversed x-axis in ggplot2 required careful use of scale_x_log10() and scale_x_reverse().

iii. Plot Aesthetics: Matching the original font size, label positions, and color intensity was time-consuming. For example, label overlap required adjusting geom_text() positioning.

## Outcome

The final scatter plot accurately captured the intended economic narrative from the original article. It successfully showed the inverse relationship between frequency and pricing level — aligning with the hypothesis that algorithmic retailers both price higher and update more often. This graph helped validate the log premium values from the regression table and served as a bridge between graphical and tabular analysis.

Merged data was used to produce a scatter plot as displayed in the code below:

```{r, eval=FALSE, echo=TRUE}
library(ggplot2)

ggplot(plot_data, aes(x = hour_gap, y = price_index, label = website)) +
  geom_point(size = 3, color = "steelblue") +
  geom_text(nudge_y = 2, size = 3.5) +
  labs(
    x = "Median Hour Gap Between Price Changes",
    y = "Price Index"
  ) +
  theme_minimal(base_size = 12)
```

------------------------------------------------------------------------

# Publishing the Quarto Presentation on GitHub

After completing the Quarto .qmd presentation in RStudio, I published it as a live HTML webpage using GitHub Pages. The process involved the following key steps:

1.  I saved my Quarto presentation (index.qmd) and all related files — including images and outputs — in a folder structured to support Quarto rendering. All image paths were stored inside a docs/images/ folder.

2.  I used RStudio’s Render button to generate the HTML output. This created an index.html file in the docs/ folder, which GitHub Pages requires to display the website.

3. I configured the git and github repository, by pushing the updated folder to GitHub. In my GitHub repository, I went to Settings \> Pages, selected the main branch and the /docs folder as the source, and saved it. GitHub then generated a public URL, **https://nanaosei95.github.io/Replica/** , which I shared with others.

4. Fixing Broken Images: Initially, images did not load on other devices because absolute paths were used (e.g., C:/Users/...). This was resolved by Converting all image references to relative paths, then ensuring that image files were tracked with Git (git add docs/images/...). I then committed and pushed the updated files to GitHub again. This process made the presentation publicly accessible and shareable.

# Making the Pull Request

A mandatory task was to create a pull request to demonstrate basic skill in GitHub collaboration workflows. Chevis Mbatha (2023) provides an accessible process of the pull request workflow, [@mbatha2023github]. I successfully executed this following the steps outlined below:

1.  Fork the repository: I first navigated to the public [repository](https://github.com/hubchev/make_a_pull_request) and forked it to my personal GitHub account.

2.  Clone the forked repository: I cloned the forked repository to my local system via the git clone command in Rstudio.

3.  Create a new branch: I created a separate branch using "git checkout -b add Tony-Osei" to safely make changes without affecting the main branch.

4.  Add my name to the Contributors.md file: I used Rstudio to open the file and added my name and title of the paper to the contributors list as instructed.

5.  Commit changes: I staged the modified file and committed the change with a meaningful message using git add and git commit.

6.  Push to GitHub: I pushed the new branch to my GitHub repository using "git push -u origin add-Tony-Osei".

7.  Create a pull request: Finally, I opened the original repository, clicked “Compare & Pull Request,” and submitted the pull request for review.

# Challenges Encountered

Apart from the problems i have listed earlier, I faced several broader challenges related to workflow, version control, and tooling within RStudio and GitHub. Below are the most notable ones:

## Publishing the Project on GitHub

A major issue was pushing the project to GitHub due to the analysis_data.dta file exceeding GitHub’s 100MB limit. Initial upload attempts failed. To resolve this, a .gitignore file was added to exclude large and temporary files (e.g., .RData, index.html). Despite this, hidden large files still caused push errors. Git Large File Storage (LFS) was considered but skipped due to setup complexity. Instead, the BFG Repo-Cleaner was used to clean Git history, requiring Java and terminal commands. This GitHub fix took over an hour and was completed with help from a teammate (Christian) just before the final deadline.

## Quarto Rendering and Visual Mode Issues

Quarto’s visual mode caused R code chunks to disappear or render as plain text. This forced all editing to be done in source mode. Switching views risked formatting loss. Additionally, some chunks failed during rendering due to undefined objects. Setting eval = FALSE ensured code displayed correctly without executing — preserving reproducibility and avoiding runtime errors.

# Collaboration and Experience

This project was completed in a group of three, though active collaboration occurred only between Christian and me. Despite repeated attempts to involve the third member, he did not participate. Christian and I divided tasks based on selected graphs and tables, regularly consulting each other to troubleshoot code, resolve formatting and rendering issues, and coordinate the GitHub deployment. Our consistent collaboration ensured the project stayed on track and maintained a high standard of accuracy and presentation.

## Utilisation of ChatGpt

 I would like to emphasize that, a key factor in the successful completion of this project was the consistent use of ChatGPT as an interactive coding assistant. With no prior experience in R or Quarto, I relied heavily on ChatGPT to:

i.  Translate complex Stata .do files into R code, preserving the logic of fixed effects models, time-series transformations, and graphical syntax.

ii. Debug errors and interpret cryptic messages that appeared during rendering, GitHub syncing, or dataset operations.

iii. Design reproducible graph code that matched the figures in the original paper.

iv. Simplify technical explanations, which helped me learn and understand the reasoning behind each step rather than blindly executing code.

v.  Create a clean report structure, including Quarto syntax, YAML headers, chunk options, and reference integration.

## Reflections on Learning RStudio as a Beginner

This project was my first hands-on experience with RStudio and served as a valuable introduction to data analysis. With no prior background in R, I learned to navigate the RStudio interface, write structured code using packages like dplyr, ggplot2, and fixest, and work efficiently with large datasets. I also gained experience rendering documents with Quarto and resolving common coding errors. More importantly, the project strengthened my problem-solving skills and taught me how to break down and document complex analytical tasks — skills I can now apply in both academic and professional contexts.

# Conclusion

The project successfully replicated core empirical results from Competition in Pricing Algorithms using RStudio. This required careful data wrangling, econometric modeling, and graph construction. Beyond replication, the process reinforced valuable skills in reproducible research and collaborative coding. The graphical replications closely matched those in the original paper, and the replication revealed the importance of both algorithmic pricing and its timing across online retailers. The project demonstrated how subtle differences in pricing strategy can be uncovered through detailed empirical work.

## What could have been accomplished with more time

If more time had been available, I could have replicated a broader set of figures and tables. I planned to replicate a regression table, which i eventually aborted. Additional effort could have gone into improving visual design to match the original formatting more precisely, that is,  combining graphs into panels, and refining x-axis alignment, or reproducing LaTeX-formatted tables. Also, i could have made my R code more flexible and reusable so that it can be applied to other datasets easily in the future. Lastly, writing a script that performs multiple steps in the replication process like, cleaning, transforming and graphing automatically, would have saved time and produced less errors. 

## Further investigations and readings

Looking ahead, several avenues for further investigation remain. Future work could involve extending the analysis to additional time periods, incorporating new retailers, or exploring how algorithmic pricing interacts with promotional campaigns and consumer demand elasticity. Moreover, the evolution of pricing algorithms under competitive pressure—especially with the rise of machine learning-based systems—warrants continued empirical attention. Researchers may also benefit from engaging with recent literature on algorithmic collusion, real-time pricing strategies, and the ethics of automated pricing to deepen the interpretive lens. Replication efforts like this one offer a foundation for such expanded inquiries, contributing to more transparent, accountable, and dynamic understanding of digital marketplaces.

------------------------------------------------------------------------

# References

<!-- References will auto-populate in the refs div below -->

::: {#refs}
:::

# This Section Is an Appendix {#apx-a}

## Relevant Links

[Link to the presentation](https://nanaosei95.github.io/replicapresentation/)

[Link to the presentation repository](https://github.com/Nanaosei95/replicapresentation.git)

[Link to the replication repository](https://github.com/Nanaosei95/Replica.git)

# Affidavit

I hereby affirm that this submitted paper was authored unaided and solely by me. Additionally, no other sources than those in the reference list were used. Parts of this paper, including tables and figures, that have been taken either verbatim or analogously from other works have in each case been properly cited with regard to their origin and authorship. This paper either in parts or in its entirety, be it in the same or similar form, has not been submitted to any other examination board and has not been published.

I acknowledge that the university may use plagiarism detection software to check my thesis. I agree to cooperate with any investigation of suspected plagiarism and to provide any additional information or evidence requested by the university.

The report includes:

-   [x] About 4000 words per student (+/- 500).

-   [x] The submission contains the Quarto file of the report.

-   [x] The submission contains the Quarto file of the presentation.

-   [x] The submission contains the HTML file of the report.

-   [x] The submission contains the HTML file of the presentation.

-   [x] The submission contains the PDF file of the report.

-   [x] The submission contains the PDF file of the presentation.

-   [x] The title page of the presentation and the report contain personal details (name, email, matriculation number).

-   [x] The report contains an abstract.

-   [x] The presentation and the report contain a bibliography, created using BibTeX with APA citation style.

-   [x] The report contains R code that proofs the student(s) expertise in coding.

-   [x] The report includes an introduction to guide the reader and a conclusion summarizing the work and discussing potential further investigations and readings, respectively.

-   [x] All significant resources used in the report and R code development.

-   [x] The filled out Affidavit.

-   [x] A concise description of the successful use of Git and GitHub, as detailed here: <https://github.com/hubchev/make_a_pull_request>.

-   [x] The link to the presentation and the handout published on GitHub.

    Osei Tony Attefah

    18.07.2025

    Koln, Germany
